\MakeUppercase{Linear versus centred Colouring Numbers}\thanks{This research was partly funded by NSERC.}

### 2022-03-31 lin-vs-cen.tex
Finally, an approach that's going to work

\setlength{\parskip}{1ex}

@@

\title{\MakeUppercase{Linear versus centred Colouring Numbers}\thanks{This research was partly funded by NSERC.}}
\author{Prosenjit Bose, Vida Dujmović, Mehrnoosh Javarsineh, and Pat Morin}

@@

\begin{thm}[\citet{kun.obrien.ea:polynomial}]\label{kun_obrien_general}
  For any graph $G$, $\chicen(G)\le \chilin^{190}(G)\cdot\log^{O(1)}(\chilin(G))$.
\end{thm}

@@

\subsection{Subdivided Grids}

Refer to \cref{grid_figure}.  For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.  For each $i\in\{1,\ldots,a\}$, the \emph{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.  We define the \emph{boundary} of $G_{a\times b}$ as $\partial G_{a\times b}:=\{1,a\}\times\{1,\ldots,b\}\cup \{1,\ldots,a\}\times\{1,b\}$.  We say that a path $P:=v_0,v_1,\ldots,v_\ell$ in $G_{a\times b}$ \emph{goes straight} at a vertex $v_i$, $i\in\{1,\ldots,k-1\}$ if $v_{i-1}$, $v_i$, and $v_{i+1}$ are contained in a single row or column.  Otherwise, we say that $P$ \emph{bends} at $v_i$.

\begin{figure}
  \begin{center}
    \begin{tabular}{c@{\hspace{2cm}}c}
      \includegraphics{figs/grid-1} &
      \includegraphics{figs/grid-2}
    \end{tabular}
  \end{center}
  \caption{The grid $G_{10}$ and its $1$-subdivision $G^{\onesub}_{10}$.}\
  \label{grid_figure}
\end{figure}

% An $a'\times b'$ \emph{subgrid} of $G_{a\times b}$ is an induced graph of the form $G_{a\times b}[\{(v,w): v\in\{i,\ldots,i+a'-1\},\, w\in\{j,\ldots,j+b'-1\}$ for some $i\in\{1,\ldots,a-a'+1\}$ and $j\in\{1,\ldots,b-b'+1\}$.   The \emph{vertex boundary} of $G_{a\times b}$ is $\{1,a\}\times\{1,\ldots,b\}\cup \{1,\ldots,a\}\times\{1,b\}$.  The \emph{edge boundary} of $G_{a,b}$ consists of all edges incident to at least one vertex in the vertex boundary of $G_{a,b}$. The \emph{boundary} of $G_{a\times b}$ is the union of the edge and vertex boundaries.

For brevity, we use $G_{k}:=G_{k\times k}$ as a shorthand for the $k\times k$ grid.  We will work extensively with the $1$-subdivision $G^{\onesub}_k$ of $G_k$.  In this case, we define $\partial G^{\onesub}_k:= \partial G_k\cup N_{G^{\onesub}_k}[\partial G_k]$.\todo{Define open and closed neighbourhods $N_G(\cdot)$ and $N_G[\cdot]$.}

\begin{lem}\label{seven_by_seven}
  Let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{7})\setminus \partial G^{\onesub}_7$ with $|I\cup L\cup Q|\le 3$ that have the following properties:
  \begin{compactenum}[(i)]\setcounter{enumi}{0}
    \item Each vertex is $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
    \item No vertex in $V(G^{\onesub}_{10})$ is adjacent to three vertices of $Q$.
    \item No vertex in $I\cup L$ is adjacent to two vertices of $Q$.
  \end{compactenum}
  Then for any pair of boundary vertices $s$ and $t$, $G_{10}$ contains a path $P$ that
  \begin{compactenum}[(a)]
    \item begins at $s$ and ends at $t$;
    \item contains every vertex in $I\cup L\cup Q$;
    \item goes straight through each vertex in $I$; and
    \item bends at each vertex in $L$.
    % \item contains no boundary vertices or edges except $s$ and $t$.
  \end{compactenum}
\end{lem}

\begin{proof}
  Case analysis(?).
\end{proof}

It will be helpful to take some natural definitions for $G_k$ and extend them to $G^{\onesub}_k$.  To do this, we associate each vertex $v$ of $G^{\onesub}_k$ with a vertex $v^{\leftup}$ of $G_k$. For each vertex $v\in V(G_k)$, we define $v^{\leftup}:=v$. For each subdivision vertex $s$ in $G^{\onesub}_k$ with neighbours $v$ and $w$, we define $s^{\leftup}$ to be the lexicographically smaller of $v$ or $w$. In other words, $s^{\leftup}$ is the vertex of $G_k$ that is directly below or the left of $s$.

For two vertices $v_1,v_2\in V(G^{\onesub}_k)$, with $v^{\leftup}_1:=(x_1,y_1)$ and $v^{\leftup}_2:=(x_2,y_2)$ we define the $L_\infty$-distance between $v_1$ and $v_2$ as $\|v_1v_2\|_\infty = \max\{|x_1-x_2|, |y_1-y_2|\}$. Refer to \cref{ball_figure}.  For each $v\in V(G^{\onesub}_k)$, the \emph{($L_\infty$) ball} of radius $\beta$ centered at $v$ is $B_{\beta}(v):=\{w\in V(G^{\onesub}_k):\|vw\|_\infty \le \beta\}$.  Because it comes up in several places, it is worth noting that $|B_{\beta}(v)|\le 4(\beta+1)^2$, and this bound is tight for each $v\in\{\beta+1,\ldots,\beta-r-1\}\times\{\beta+1,\ldots,b-\beta-1\}$.

\begin{figure}
  \begin{center}
    \includegraphics{figs/grid-3}
  \end{center}
  \caption{The ball $B_2(v)=B_2(s)$ in $G^{\onesub}_{10}$.}
  \label{ball_figure}
\end{figure}

\begin{lem}\label{four_points}
  Let $P$ be a set of at most $3$ vertices of $G^{\onesub}_k$.  Then there exists a set $S$ of pairwise-disjoint balls, each of radius $3$, such that each vertex in $P$ is contained in the interior of some ball in $S$.\todo{Define the interior of a ball.}
\end{lem}

\begin{proof}
  Let $P:=\{v_1,\ldots,v_t\}$ where $p_i^{\leftup}:=(x_i,y_i)$.  Let $X:=\bigcup_{i=1}^t \{x_i\bmod 7,(x_{i+1})\bmod 7\}$ and let $Y:=\bigcup_{i=1}^t\{y_i\bmod 7,(y_i+1)\bmod 7\}$.  Then $|X|,|Y|\le 2t\le 6$.  In particular, there exists some $x\in\{0,\ldots,6\}\setminus X$ and some $y\in\{0,\ldots,6\}\setminus Y$. For each $i,j\in\Z$, let $c_{i,j}:=(7i+x+3,7j+y+3)$ and observe that the ball $B_{3}(c_{i,j})$ does not have any point of $P$ on its boundary.  Furthermore, the set of balls $S:=\{B_{3}(c_{i,j}):i,j\in\Z\}$ is a partition of $V(G^{\onesub}_{k})$.  Therefore $S$ is a set of pairwise-disjoint balls, each of radius $3$, such that each vertex in $P$ is contained in the interior of some ball in $S$.
\end{proof}

\begin{lem}
  Let $P$ be a subset of $V(G^{\onesub}_k)$ such that no ball of radius $18$ contains more than $3$ vertices in $P$.  Then there exists a set $S$ of disjoint balls, each of radius $3$ and such that each point of $P$ is in the interior of some ball in $S$.
\end{lem}

\begin{proof}
  Define a graph $H$ with vertex set $V(H):=P$ and that contains an edge $pq$ if and only if $\|pq\|_\infty \le 12$. Observe that, if some connected subgraph of $H$ contains $4$ vertices, then these $4$ vertices are contained in a ball of radius $18$.  Therefore each component of $H$ contains at most $3$ vertices.

  For each component $C$ of $H$, apply \cref{four_points} to $V(C)$ and keep only those balls in $S$ that contain at least one vertex of $V(C)$.  This produces a set $S_C$ of disjoint balls, each of radius $3$, that each contain some vertex in $V(C)$ in their interior.  Any two balls $B_{C_1}$ and $B_{C_2}$ obtained from different components of $H$ are disjoint because the first ball contains a vertex $p_1\in P$ and the second ball contains a vertex $p_2\in P$ such that $\|p_1p_2\|\infty > 12$, but the balls $s_1$ and $s_2$ each have radius $3$.
\end{proof}

\begin{lem}\label{one_path_enchilada}
  Let $k$ be a positive integer, let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{k})\setminus\partial G^{\onesub}_k$ that have the following properties:
  \begin{compactenum}[(1)]\setcounter{enumi}{-1}
    \item Any ball of radius 18 contains at most $3$ elements of $I\cup L\cup Q$.
    \item Each vertex in $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
    \item No vertex in $V(G^{\onesub}_{10})$ is adjacent to three vertices of $Q$.
    \item No vertex in $I\cup L$ is adjacent to two vertices of $Q$.
  \end{compactenum}
  Then $G^{\onesub}_k$ contains a path $P$ that
  \begin{compactenum}[(a)]
    \item contains all the vertices in $I\cup L\cup Q$;
    \item goes straight through each vertex in $I$; and
    \item turns at each vertex in $L$.
  \end{compactenum}
\end{lem}

\begin{proof}
  Start with the snakelike path $P_0$ that visits every $7$th row of $G_k$. Cover the elements of $I\cup L\cup Q$ with disjoint balls of side length $3$.  Each such ball is intersected by exactly one row of the snakelike path.  Use \cref{seven_by_seven} to locally deform $P_0$ to pick up the objects in each ball that it intersects.
\end{proof}

\subsection{Approximate Voronoi Diagrams in $G^{\onesub}_k$}

% \todo[inline]{This lemma is not true. Pick a center $x$ and place a point in each of the quadrants around $x$ where the $i$th point is at distance $10\beta - (ir)$ from $x$.  Then $x$ is incident on all four Voronoi regions.  We really have to use the fact that $\beta$ (or rather, $\beta+ar$) is also an upper bound on the distance we actually care about.}
\begin{lem}\label{voronoi_mapping}
  Let $\beta, k\in\N$, let $P\subseteq V(G^{\onesub}_k)$ be such that $\min\{\|pq\|_\infty:p,q\in\binom{P}{2}\} \ge \beta$, and define  $Q:=\bigcup_{p\in P} B_\beta(p)$. Then there exists a function $\vor:Q\to P$ such that
  \begin{compactenum}[(i)]
    \item \label{real_close} For each $p\in P$ and each $v\in B_{\beta/8-1}(p)$, $\vor(v)=p$.
    \item \label{never_far} For each $v\in Q$, $\|v\vor(p)\|_\infty \le 7\beta/4$.
    \item \label{three_in_ball} For any ball $B$ of radius $\beta/12$, $|\{\vor(v):v\in B\cap Q\}|\le 3$.
    \item \label{twocolour_neighbourhood} For any $v\in Q$, $|\{\vor(v)\}\cup\{\vor(w):w\in Q\cap N_{G^{\onesub}_k}(v)\}|\le 2$.
  \end{compactenum}
\end{lem}

\begin{proof}
  We first define a function $\vor':Q\to P$ that satisfies (\ref{real_close})--(\ref{three_in_ball}) and then show how local modifications can be used to obtain a function $\vor$ that also satisfies (\ref{twocolour_neighbourhood}).

  Arbitrarily order the vertices of $P$ as $p_1,\ldots,p_{|P|}$.  For each $v\in Q$, let $d(v):=\min\{\|vp\|_\infty:p\in P\}$, let $S(v):=\{p\in P:\|vp\|_\infty \le d(v)+3\beta/4\}$, let $i(v):=\min\{i:p_i\in S(v)\}$, and let $\vor'(v)=p_{i(v)}$.  In words, $d(v)$ represents the distance from $v$ to its nearest neighbour in $P$, $S(v)$ represents the set of approximate nearest neighbours of $v$ in $P$ (up to an additive error of $\beta/12$), and $\vor'(v)$ is the approximate nearest neighbour of $v$ of smallest index.

  To see that $\vor'$ satisfies (\ref{real_close}), observe that, for each distinct $p,q\in P$ and $v\in B_{\beta/8-1}(p)$, the triangle inequality implies that $\|vq\|_\infty \ge \|pq\|_\infty-\|vp\|_\infty \ge \beta-\beta/8+1 > 7\beta/8$.  On the other hand $d(v)\le \beta/8$, so $d(v)+3\beta/4\le 7\beta/8$.  Therefore, $q\not\in S(v)$ for any $q\in P\setminus\{p\}$, so $\vor'(v)=p$.

  That $\vor'$ satsifies (\ref{never_far}) follows immediately from the fact that each $v\in Q$, so $d(v)\le \beta$, and therefore $\|pv\|_\infty\le 7\beta/4$ for each $p\in S(v)$.

  We now argue that $\vor'$ satisfies (\ref{three_in_ball}).  Let $v_1,\ldots,v_4\in Q$ and $1\le i_1<\cdots<i_4\le|P|$ be such that $\vor'(v_j)=p_{i_j}$ for each $j\in\{1,\ldots,4\}$.  Suppose, for the sake of contradiction, that $i_1,\ldots,i_4$ are all distinct and that $v_1,\ldots,v_4$ are contained in some ball $B_{\beta/12}(c)$.  By the triangle inequality $\|v_iv_j\|_\infty \le \|v_ic\|_\infty+\|cv_j\|_\infty \le \beta/6$ for each  $i,j\in\{1,\ldots,4\}$.
  Since $v_1\in Q$, $d(v)\le\beta$ so
  \[
    \|p_{i_1}v_1\|_\infty \le d(v_1) + 3\beta/4 \le \beta + 3\beta/4 \enspace .
  \]
  Since $\vor'(v_2)=p_{i_2}\neq p_{i_1}$,
  \[
    \|p_{i_2}v_2\|_\infty
     < \|p_{i_1}v_2\|_\infty - 3\beta/4
     \le \|p_{i_1}v_1\|_\infty + \beta/6 - 3\beta/4
     \le \beta + \beta/5 \enspace .
  \]
  Since $\vor'(v_3)=p_{i_3}\neq p_{i_2}$,
  \[
     \|p_{i_2}v_3\|_\infty
     < \|p_{i_2}v_3\|_\infty - 3\beta/4
     \le \|p_{i_2}v_2\|_\infty + \beta/6 - 3\beta/4
     < \beta + 2\beta/6 - 3\beta/4 \enspace .
  \]
  Finally, since $\vor'(v_4)=p_{i_4}\neq p_{i_3}$,
  \[
    \|p_{i_4}v_4\|_\infty
    < \|p_{i_3}v_4\|_\infty - 3\beta/4
    \le \|p_{i_3}v_3\|_\infty + \beta/5 - 3\beta/4
    < \beta + 3\beta/6 - 3\beta/2 = 0 \enspace ,
  \]
  which is clearly a contradiction.  This establishes that $\vor'$ satisfies Condition~(\ref{three_in_ball}).\todo{TODO: Optimize and cleanup. Finish proving (iii)}
\end{proof}

\subsection{Pseudogrids}

A graph $H$ is an $a\times b$ \emph{pseudogrid} if there exists a set $\mathcal{P}$ of vertex-disjoint paths in $H$ such that the quotient graph $H':=H/\mathcal{P}$ is isomorphic to a subdivision of $G_{a\times b}$.  The following lemma connecting grid minors and pseudogrids is stated implicitly by \citet{kun.obrien.ea:polynomial}.  We give a proof here because some of the details of the connection are required in order to prove our results.

\begin{lem}\label{pseudogrid_minor}
  If a graph $G$ contains a $G_{a\times b}$ grid minor then $G$ contain a subgraph $H$ that is an $a\times b$ pseudogrid.
\end{lem}

\begin{proof}
  (This proof only uses the fact that $G_{a\times b}$ has maximum degree  $4$.)  Since $G$ contains a $G_{a\times b}$ minor then, by definition, $G$ contains a subgraph $H$ and a partition $\mathcal{B}:=(B_x:x\in V(G_{a\times b}))\}$ of $V(H)$ such that
  \begin{inparaenum}[(i)]
    \item $B_x$ is non-empty and $H[B_x]$ is connected for each $x\in V(G_{a\times b})$;
    \item $H/\mathcal{B}$ is isomorphic to $G_{a\times b}$ using the isomorphism given by $B_x\mapsto x$.
  \end{inparaenum}
  Let $(H,\mathcal{B})$ be chosen so that $|E(H)|$ is minimum, over all choices of $(H,\mathcal{B})$ that satisfy (i) and (ii).

>>>>

\newcommand{\defin}[1]{\emph{\color{brown}#1}}
\setlength{\parskip}{1ex}

@@

\title{\MakeUppercase{Linear versus centred Colouring Numbers}\thanks{This research was partly funded by NSERC.}}
\author{Prosenjit~Bose, Vida~Dujmović, Hussein~Houdrouge, Mehrnoosh~Javarsineh, and Pat~Morin}

\DeclareMathOperator{\VE}{VE}

@@

\begin{thm}[\citet{kun.obrien.ea:polynomial}]\label{kun_obrien_general}
  For any graph $G$, $\chicen(G)\le \chilin^{190}(G)\cdot\log^{O(1)}(\chilin(G))$.\todo{Update with most recent numbers}
\end{thm}

@@

Refer to \cref{grid_figure}.  For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.  For each $i\in\{1,\ldots,a\}$, the \emph{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.

In this work, we need to work with pseudogrids.  An \defin{$a\times b$-pseudogrid} is any graph that can be obtained from $G_{a\times b}$ in the following way:
\begin{compactenum}
  \item Replace each edge $vw$ of $G_{a\times b}$ with a path $\overline{P}_{vw}$ whose endpoints are $v$ and $w$.  (In other words, $\overline{P}_{vw}$ is a path obtained by subdividing the edge $vw$ zero or more times.)
  \item Replace each degree-$4$ vertex $(x,y)$ of $G_{a\times b}$ with a path $P_v$. If $P_v$ has only one vertex then this operation has no effect.  Otherwise, $P_v$ has two endpoints $p$ and $q$ and
  \begin{compactenum}[(Q1)]
    \item $p$ is adjacent to $\{(x-1,y), (x,y-1)\}$ and $q$ is adjacent to $(x+1,y),(x,y+1)$;
    \item $p$ is adjacent to $\{(x,y+1), (x-1,y)\}$ and $q$ is adjacent to $(x,y-1),(x+1,y)$; or
    \item $p$ is adjacent to $\{(x-1,y),(x+1,y)\}$ and $q$ is adjacent to $\{(x,y-1),(x,y+1)\}$.\label{q_iii}
  \end{compactenum}
\end{compactenum}

Let $G$ be an $a\times b$ pseudogrid.  For an edge $vw$ of $G_{a\times b}$, we let $P_{vw}:=\overline{P}_{vw}-\{v,w\}$ denote the (possibly empty) subpath containing the internal vertices of $\overline{P}_{vw}$.  For each vertex $v$ of $G_{a\times b}$ of degree less than $4$ we define $P_{v}$ to be the $1$-vertex path that contains only $v$.  In this way, $\mathcal{P}:=\{V(P_v):v\in V(G_{a\times b})\}\cup\{V(P_{vw}):vw\in E(G_{a\times b})\}$ is a partition of $V(G)$ into induced paths.  We call $\mathcal{P}$ the \defin{grid-partition} of $G$ and we denote it by $(P_\mu:\mu\in V(G_{a\times b})\cup E(G_{a\times b}))$.  This defines a mapping $\delta:V(G)\to V(G_{a\times b})\cup E(G_{a\times b})$ where $u\in V(P_{\delta(u)})$ for each $u\in v(G)$.

% Then each vertex $u$ of $G$ naturally corresponds to an edge or vertex $u'$ of $G_{a\times b}$:  If $u$ is a vertex of $P_{vw}$ then $u'=vw$; if $u$ a vertex of $P_v$ then $u'=v$; and otherwise $u'=u$.  This latter case occurs precisely when $u'$ is one of the boundary vertices of $G_{a\times b}$.

Each row $R':=v_1,\ldots,v_a$ of $G_{a\times b}$ corresponds naturally to a path $R$ of $G$. The path $R$ contains $V(\overline{P}_{v_iv_{i+1}})$ for each $i\in\{1,\ldots,a-1\}$.  However, for $i\in\{2,\ldots,a-1\}$ $R$ may or may not contain $V(P_{v_i})$.  In particular, if $P_{v_i}$ was created using (Q\ref{q_iii}) then $R$ does not contain any internal vertices in $P_{v_i}$.
Similarly, a column $C':=v_1,\ldots,v_b$ of $G_{a\times b}$ corresponds to a path that contains $V(\overline{P}_{v_jv_{j+1}})$ for each $j\in\{1,\ldots,b-1\}$. This correspondence allows us to talk about the rows and columns of $G$, which we will do next.

As part of our proof, we use the operation of \defin{deleting} a row (or column) $R$ of $G$, which works as follows:  We remove all edges of $R$ from $G$ and eliminate any isolated vertices.  If this produces vertices of degree $1$ (which happens when $R$ is the first or last row of $G$) then we repeatedly remove degree-$1$ vertices until none remain.  If $G$ is an $a\times b$ pseudogrid and we delete some row $R$, then the resulting graph is an $a\times (b-1)$ pseudogrid.  Similarly, if we delete column $C$ of $G$, then the resulting graph is an $(a-1)\times b$ pseudogrid.

\subsection{Pseudogrids with Exclusively Frequent Colours}

Let $G$ be a $k\times k$ pseudogrid with grid partition $\mathcal{P}:=(P_\mu:\mu\in\VE(G_{k\times k}))$ and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$.  For any $\mu\in\VE(G_{k\times k})$, we let $\varphi_{\mathcal{P}}(\mu):=\{\phi(v):v\in V(P_\mu)\}$.  For any colour $\alpha\in\{1,\ldots,c\}$, define $\varphi^{-1}(\alpha):=\{v\in V(G):\varphi(v)=\alpha\}$ and define $\varphi_\mathcal{P}^{-2}(\alpha):=\{\mu\in \VE(G_{k\times k}):\alpha\in\varphi_\mathcal{P}(\mu)$.  For any colour set $A\subseteq\{1,\ldots,c\}$ define $\varphi^{-1}(A):=\bigcup_{\alpha\in A}\varphi^{-1}(\alpha)$ and $\varphi_\mathcal{P}^{-2}(A):=\bigcup_{\alpha\in A}\varphi_{\mathcal{P}}^{-2}(\alpha)$.

% We extend $\varphi$ to a partial colouring $\phi:V(G_{k\times k})\cup E(G_{k\times k})\to\{1,\ldots,c,\perp\}$ of the edges and vertices of $G_{k\times k}$ as follows:  For each vertex $v$ of $G_{k\times k}$, define $\phi(v)$ be an arbitrary colour in $\{\varphi(x):x\in V(P_v)\}$.  For each edge $vw$ of $G_{k\times k}$, there are two possibilities:
% If $P_{vw}$ is empty then $\phi(vw):=\perp$, otherwise $\phi(vw)$ is an arbitrary colour in $\{\varphi(x):x\in V(P_{vw})\}$.

\begin{lem}
  Let $0 < \epsilon < 1$, let $k\ge 1/\epsilon$ be an integer, let $G$ be a $k\times k$ pseudogrid and let $\varphi:V(G)\to\{1,\ldots,\lfloor\epsilon k\floor\}$ be a vertex colouring of $G$.
  Then $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ that has a grid-partition $\mathcal{P}':=(P'_\mu:\mu\in \VE(G_{k'\times k'})$ such that
  for any $A\subseteq \varphi(V(G'))$, $|\varphi_{\mathcal{P}'}^{-2}(A)| \ge |A|/2\epsilon$.
\end{lem}

\begin{proof}
  The proof is by induction on $|\varphi(V(G))|$, the number of colours used by the colouring $\varphi$.  If there exists no $A\subseteq\{1,\ldots,\lfloor\epsilon k\rfloor\}$ with $\varphi_{\mathcal{P}}^{-2}(A) < |A|/2\epsilon$ then taking $G':=G$ and $\mathcal{P}':=\mathcal{P}$ satisfies the requirements of the lemma. Now assume that such a set $A$ exists.

  We will remove a set $R$ of rows and a set $C$ of columns from $G$ with $|R|=|C|\le |A|/2\epsilon$.  For each $v:=(i,j)\in V(G_{k\times k})$ with $A\cap\varphi(v)\neq\emptyset$, we include row $j$ in $R$ and column $i$ in $C$.  For each edge $vw\in E(G_{k\times k})$ with $A\cap\phi()\alpha\in\Phi(vw)$, we include the row of $G$ that contains $P_{vw}$ in $R$ or we include the column of $G$ that contains $P_{vw}$ in $C$.  Finally, we add arbitrary rows to $R$ or columns to $C$ to ensure that $|R|=|C|$.  Now we remove all rows in $R$ and all columns in $C$ from $G$ to obtain a $k_0\times k_0$ pseudogrid $G_0$ with $k_0\ge k-|A|/2\epsilon$ and such that $\varphi(V(G_0))\cap A=\emptyset$.  In particular, $|\varphi(V(G_0))|\le |\varphi(V(G))|-|A|$.

  \todo[inline]{Now apply induction......}
\end{proof}

\begin{lem}
  Let $C>1$ be an integer, let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}$, and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a vertex colouring of $G$ such that, for any $A\subseteq\varphi(G)$, $|\varphi_{\mathcal{P}'}^{-2}(A)| \ge C\cdot|A|+1$. Then there exists a partial colouring $\phi:\VE(G_{k\times k})\to\{1,\ldots,c,\perp\}$ with the following properties:
  \begin{compactenum}[(i)]
    \item For each $\mu\in\VE(G_{k\times k})$, $\phi(\mu)=\perp$ or $\phi(\mu)\in\varphi(\mu)$.
    \item For each $\alpha\in\varphi(V(G))$, $|\{\mu\in\VE(G_{k\times k}):\phi(\mu)=\alpha\}|\ge C$.
  \end{compactenum}
\end{lem}

\begin{proof}
  Use Hall's Theorem to find $C$ matchings such that each colour $\alpha$ appears in all $C$ matchings and each element of $\VE(G_{k\times k})$ appears in at most one matching.
\end{proof}

\subsection{Finding Paths Through Well-Separated Vertex/Edge Pairs}

The \defin{$r$-box} centered at a vertex $v:=(i,j)$ of $G_{a\times b}$ is defined as
\[  B_r(v) := \{i-r,\ldots,i+r\}\times\{j-r,\ldots,j+r\} \cap V(G_{a,b}) \enspace . \]
The \defin{$r$-box} centered at an edge $vw$ of $G_{a\times b}$ is $B_r(vw):=B_r(v)\cup B_r(w)$.  Observe that, for any $\mu\in V(G_{a\times b})\cup E(G_{a\times b})$,  $v\in B_r(u)$ if and only if $u\in B_r(v)$.  Any $r$-box $B_r(\mu)$ defines an induced subgraph that we denote by $G_r(\mu):=G_{a\times b}[B_r(\mu)]$.  We extend these definitions to vertices of an $a\times b$ pseudogrid $G$ so that
\[
   \tilde{B}_r(v) := \bigcup_{\mu\in V(G_r(\delta(v)) \cup E(G_r(\delta(v))} V(P_\mu)
\]
and $\tilde{G}_r(v):=G[\tilde{B}_r(v)]$.

\begin{lem}\label{pick_up_two}
  Let $v$ be any vertex of $G$, let $w$ be any vertex of $\tilde{G}_r(v)$, let $s$ be any vertex in the leftmost column of $\tilde{G}_r(v)$ and let $t$ be any vertex in the rightmost column of $\tilde{G}_r(v)$.  Then there exists an $st$-path\todo{define $st$-path} in $G_r(v)$ that contains $v$ and $w$.
\end{lem}

\begin{proof}
  Let $i_s \le i_v \le i_w \le i_t$ be the columns of $\tilde{G}_r(v)$ that contain $s$, $v$, $w$, and $t$, respectively.  Note that the assumption that $i_v \le i_w$ is without loss of generality and that $i_s \le i_t - r -1$.\todo{finish up}
\end{proof}

\begin{lem}
  Let $r$ be an integer and let $S\subseteq V(G)$ be such that $|\tilde{B}_{r}(v)\cap S|=|\tilde{B}_{2r}\cap S|\le 2$ for each $v\in S$.  Then $G$ contains a path that contains every vertex in $S$. \todo{Set specific values for $r$}
\end{lem}

\begin{proof}[Proof sketch]
  Choose $S'\subseteq S$ so that, if $B_r(v)\cap S=\{v,w\}$ then exactly one of $v$ or $w$ is in $S'$. Now do some scan-line thing with a distance of $2r+1$ between consecutive rows.  When this scan encounters a box $B_r(v)$ for some $v\in S'$, use \cref{pick_up_two} to detour through the box to pick up the (at most two) vertices in $B_r(v)\cap S$.
\end{proof}

\subsection{Finding an Uncentered Path}

\begin{lem}
  Let $G$ be a $k\times k$ pseudogrid and let $\varphi:V(G)\to\{1,\ldots,c\}$ be a colouring of $G$ such that, for each $\alpha\in\{1,\ldots,C\}$, the set $S_\alpha:=\{\mu\in V(G_{k\times k})\cup E(G_{k\times k}): \alpha\in\Phi(\mu)\}$ has size at least $C$.  Then there exists $S\subseteq V(G_{k\times k})\cup E(G_{k\times k})$ such that
  \begin{compactenum}[(i)]
    \item \label{hits_both} $|S\cap S_\alpha|\ge 2$ for each $\alpha\in\{1,\ldots,c\}$; and
    \item \label{spread_out} $|B_r(\mu)\cap S|\le 2$ for each $\mu\in S$.
  \end{compactenum}
\end{lem}

\begin{proof}
  We construct $S$ using a two round process.  We start with an initially empty set $S_1$ and repeat the following for each $\alpha= 1,\ldots,c$:
  If there exists distinct $\mu_1,\mu_2\in S_{\alpha}$ such that
  \begin{compactenum}[(a)]
    \item $\alpha\in\Phi(\mu_1)\cap\Phi(\mu_2)$;
    \item $\mu_1\not\in B_r(\mu_2)$; and
    \item $\{\mu_1,\mu_2\}\cap \bigcup_{\mu\in S} B_{r}(\mu)=\emptyset$
  \end{compactenum}
  then include $\mu_1$ and $\mu_2$ in $S_1$ and declare this a \defin{success} for colour $\alpha$.  Otherwise, declare this a failure for colour $\alpha$.

  At the end of this process, the resulting set satisfies (\ref{spread_out}) but is only guaranteed to satisfy (\ref{hits_both}) for the successful colours $\alpha$.  We now use the second round to create a set $S_2$ to fix this. Our strategy is to use \cref{hall_vees} to find two matchings between the failed colours and the boxes in $\{B_r(\mu):\mu\in S_1\}$ that contain those colours.  Each missing colour is matched in both matching, but each box is used in at most one of the matchings.  This is always possible, since the number of edges and vertices of $G_r(\mu)$ is $\Theta(r^2)$ while the

  Consider a (for now) random permutation $\alpha_1,\ldots,\alpha_c$ of $\{1,\ldots,c\}$. For each $i\in\{1,\ldots,c\}$: Find $2(r+1)^2$ elements $\mu\in S_1$ such that
  \begin{compactenum}[(a)]
    \item
  \end{compactenum}
\end{proof}

\newpage

% \hrule
%
% \subsection{Subdivided Grids}
%
% Refer to \cref{grid_figure}.  For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.  For each $i\in\{1,\ldots,a\}$, the \emph{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.  We define the \emph{boundary} of $G_{a\times b}$ as $\partial G_{a\times b}:=\{1,a\}\times\{1,\ldots,b\}\cup \{1,\ldots,a\}\times\{1,b\}$.  We say that a path $P:=v_0,v_1,\ldots,v_\ell$ in $G_{a\times b}$ \emph{goes straight} at a vertex $v_i$, $i\in\{1,\ldots,k-1\}$ if $v_{i-1}$, $v_i$, and $v_{i+1}$ are contained in a single row or column.  Otherwise, we say that $P$ \emph{bends} at $v_i$.
%
% \begin{figure}
%   \begin{center}
%     \begin{tabular}{c@{\hspace{2cm}}c}
%       \includegraphics{figs/grid-1} &
%       \includegraphics{figs/grid-2}
%     \end{tabular}
%   \end{center}
%   \caption{The grid $G_{10}$ and its $1$-subdivision $G^{\onesub}_{10}$.}\
%   \label{grid_figure}
% \end{figure}
%
% % An $a'\times b'$ \emph{subgrid} of $G_{a\times b}$ is an induced graph of the form $G_{a\times b}[\{(v,w): v\in\{i,\ldots,i+a'-1\},\, w\in\{j,\ldots,j+b'-1\}$ for some $i\in\{1,\ldots,a-a'+1\}$ and $j\in\{1,\ldots,b-b'+1\}$.   The \emph{vertex boundary} of $G_{a\times b}$ is $\{1,a\}\times\{1,\ldots,b\}\cup \{1,\ldots,a\}\times\{1,b\}$.  The \emph{edge boundary} of $G_{a,b}$ consists of all edges incident to at least one vertex in the vertex boundary of $G_{a,b}$. The \emph{boundary} of $G_{a\times b}$ is the union of the edge and vertex boundaries.
%
%
% % Let $G^{\onesub}_k$ denote the $1$-subdivision of $G_k$.
%
% A \emph{simple $k\times k$ pseudogrid} is any graph that can be obtained from $G^{\onesub}_k$ by replacing some degree-$4$ vertices of $G_k$ with small binary trees.  Specifically, we may replace a degree-$4$ vertex with a tree whose leaves are the vertices in $N_{G^{\onesub}_k}(v)$ and whose two internal nodes $v_a$ and $v_b$ are each adjacent to two leaves and to each other.
%
% Let $G_0$ be a simple $k\times k$ pseudogrid and observe that there is a natural mapping $f:E(G)\to E(G_k)\cup V(G_k)$ which is fairly obvious except that any edge $v_av_b\in E(G_0)$ introduced when replacing a degree-$4$ vertex $v$ in $G_k$ has $f(v_av_b):= v$.
%
% \begin{lem}\label{one_path}
